---
title: "HW4_bdbarbe2"
author: "Beau Barber"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# For homework 4, please pull the folder HW4 from the class_notesrepository on GitHub. It contains the pheno.txt* document to be used in this homework. From the pheno.txt file, obtain the following information:

### Reading text file into R
```{r}
pheno <- read.table(file.choose("pheno.txt"), header = T, sep = "")
head(pheno)
```



## 1) How many unique observations are there in the “GENO” variable? (You haven’t used this function in class yet. Try to look for functions.)

```{r}
unique_obs <- unique(pheno$GENO)
length(unique_obs)
```
485 unique observations in the "GENO" variable



## 2) Save as a text file a subset of the pheno data set containing only range 1.
```{r}
pheno1 <- pheno[pheno$range == 1,]
pheno1
write.table(pheno1, file = "pheno1.txt")
```


## 3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.

Range of Y:
```{r}
range(pheno$Y, na.rm = T)
```
The range of Y is from [min = 0.5250602,max = 21.6188835]

Range of M:
```{r}
range(pheno$M, na.rm = T)
```
The range of M is from [min = 0.4502466,max = 0.8779443]

Range of h4:
```{r}
range(pheno$h4, na.rm = T)
```
The range of h4 is from [min = 1.204, max = 5.100]



## 4) What row contains the maximum value of “h4”?
```{r}
which.max(pheno$h4)
```
The 427th row of the variable h4 contains the max value of h4


## 5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).
```{r}
for (i in names(pheno[,7:12]))
  print(mean(pheno[,i], na.rm = T))
```
The means for each variable are as follows
Y is 7.99446
M is 0.7431033
h1 is 0.2766773
h2 is 1.248048
h3 is 2.621128
h4 is 3.165503.


## 6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.
```{r}
pheno2 <- pheno
pheno2$Y[pheno2$Y < 2] <- 0
pheno2
```