---
title: "Ece_Gulkirpik_HMW4"
author: "Ece Gulkirpik"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#For homework 4, please pull the folder HW4 from the class_notesrepository on GitHub. It contains the pheno.txt* document to be used in this homework. From the pheno.txt file, obtain the following information:

```{r}
pheno <- read.table("HW4/pheno.txt", header = TRUE)
pheno
```
#1) How many unique observations are there in the “GENO” variable? (You haven’t used this function in class yet. Try to look for functions.)
```{r}
geno_obs <- length(unique(pheno$GENO))
geno_obs
```
#2) Save as a text file a subset of the pheno data set containing only range 1.
```{r}
#I gave "subset_pheno" name to the subet of the pheno data that I created which included only range 1. To indicate that this subset includes only range 1, I used " == '1'" in the code.After I created the subset, I used write.table() function to save it, and added ".txt" after the file name to be able to save it as a text file. 
subset_pheno <- pheno[pheno$range == '1',]
write.table(x=subset_pheno, file= "subset_pheno.txt")
subset_pheno
```

#3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.
```{r}
#I used range() function for this question. However, there are missing values in the dataset. Therefore, I had to add "na.rm = TRUE" in the code.
range_Y <- range(pheno$Y, na.rm = TRUE)
range_M <- range(pheno$M, na.rm = TRUE)
range_h4 <- range(pheno$h4, na.rm = TRUE)
range_Y
range_M
range_h4
```
#4) What row contains the maximum value of “h4”?
```{r}
# First, I wanted to find what is the ma value of h4 variable.
max_value_h4 <- max(pheno$h4, na.rm = TRUE)
max_value_h4
# I went through the data quickly and realized that there are several rows containing the max value of h4 variable. To identify which rows contain this value I used the following code:
max_value_h4_rows <-which(pheno$h4 == max_value_h4)
max_value_h4_rows
# Finally, I wanted to see the details in these row.
pheno[c(max_value_h4_rows),]

```
#5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).
```{r}
new_pheno_data <- pheno[,7:12]
var_means <- NULL
variables <- names(new_pheno_data)
for(i in variables){
  var_means[i] <-mean(new_pheno_data[,i], na.rm=TRUE)
}
var_means
```
#6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.
```{r}
pheno2 <- pheno
pheno2$Y[pheno2$Y<2]<-0
head(pheno2)
```