---
title: "GraceTan_HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in data

```{r}
library(readr)
library(tidyverse)
setwd("~/Documents/Grad School Yr 1/R/class_notes/HW4")

colnames <- c("LOC","GENO","TAXA","range","row","BLOCK","Y","M","h1",	"h2",	"h3",	"h4")

pheno_pre <- read.table("pheno.txt", col.names=colnames)
pheno <- pheno_pre[-1, ]
view(pheno)
```

1) How many unique observations are there in the “GENO” variable?

```{r}
library(dplyr)
geno_freq <- count(pheno, vars = GENO)
(count(geno_freq))
```
*There are 485 unique observations in the "GENO" variable

2) Save as a text file a subset of the pheno data set containing only range 1

```{r}
range_1 <- pheno[which(pheno$range == 1), ]

write.table(range_1, file = "HW4_range_1.txt",
            row.names = FALSE,
            col.names = TRUE)
```

3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.

4) What row contains the maximum value of “h4”?

5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).

6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.