---
title: "GraceTan_HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in data

```{r}
library(readr)
setwd("~/Documents/Grad School Yr 1/R/class_notes/HW4")

colnames <- c("LOC","GENO","TAXA","range","row","BLOCK","Y","M","h1",	"h2",	"h3",	"h4")

pheno_pre <- read.table("pheno.txt", col.names=colnames)
pheno <- pheno_pre[-1, ]
```

1) How many unique observations are there in the “GENO” variable?

```{r}
library(dplyr)
geno_freq <- count(pheno, vars = GENO)
(count(geno_freq))
```
There are 485 unique observations in the "GENO" variable

2) Save as a text file a subset of the pheno data set containing only range 1

```{r}
range_1 <- pheno[which(pheno$range == 1), ]

write.table(range_1, file = "HW4_range_1.txt",
            row.names = FALSE,
            col.names = TRUE)
```

3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.

```{r}
range(pheno$Y, na.rm=TRUE)
range(pheno$M, na.rm=TRUE)
range(pheno$h4, na.rm=TRUE)
```

4) What row contains the maximum value of “h4”?

```{r}
max(pheno$h4, na.rm=TRUE)
pheno[which(pheno$h4 == 5.1), ]
```
Rows 428, 445, and 577 all contain the maximum value of h4, which is 5.1

5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).

First had to change from characters to numerics
```{r}
str(pheno)

Y_num <- as.numeric(pheno$Y)
M_num <- as.numeric(pheno$M)
h1_num <- as.numeric(pheno$h1)
h2_num <- as.numeric(pheno$h2)
h3_num <- as.numeric(pheno$h3)
h4_num <- as.numeric(pheno$h4)

nums <- pheno[ , 7:12]
nums[ ,1] <- Y_num
nums[ ,2] <- M_num
nums[ ,3] <- h1_num
nums[ ,4] <- h2_num
nums[ ,5] <- h3_num
nums[ ,6] <- h4_num

str(nums)

for(i in 1:6) {
  var_mean <- mean(nums[ ,i], na.rm = TRUE)
  print(var_mean)
  i <- i + 1
}
```

6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.

```{r}
pheno2 <- pheno[ ,1:12]
Y_num <- as.numeric(pheno2$Y)
pheno2[ ,7] <- Y_num

for(i in 1:1344) {
  no_value <- is.na(pheno[i,"Y"])
  if (no_value == TRUE) {
    i <- i + 1
  } else if(pheno2[i,"Y"] < 2) {
    pheno2[i,"Y"] <- 0
    print(i)
  }
  i <- i + 1
}
```
Printed values show row number of "Y" values less than 2