---
title: "Jadelyn_Stewart"
author: "Jadelyn Stewart"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Read in the text file for the homework

```{r}
pheno_1 <- read.table(file.choose("pheno.txt"), header = T, sep = "")
head(pheno_1)
```

##1) How many unique observations are there in the “GENO” variable? (You haven’t used this function in class yet. Try to look for functions.)

```{r}
uniq_value_1 <- length(unique(pheno_1$GENO))
```

##2) Save as a text file a subset of the pheno data set containing only range 1.
```{r}
pheno_2 <- pheno_1[pheno_1$range == 1,]
pheno_2
write.table(pheno_2, file = "pheno_2.txt")
```
##3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.

Range for Y:
```{r}
range(pheno_1$Y, na.rm = T)
```
Y's min = 0.5250602 and max = 21.6188835


Range for M:
```{r}
range(pheno_1$M, na.rm = T)
```
M's min = 0.4502466 and max = 0.8779443

Range for h4:
```{r}
range(pheno_1$h4, na.rm = T)
```
h4's min = 1.204 and max = 5.100

##4) What row contains the maximum value of “h4”?
```{r}
which.max(pheno_1$h4)
```
The 427th row contains the max for h4

##5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).
```{r}
for (i in names(pheno_1[,7:12]))
  print(mean(pheno_1[,i], na.rm = T))
```
The means are:
Y: 7.99446
M: 0.7431033
h1: 0.2766773
h2: 1.248048
h3: 2.621128
h4: 3.165503

##6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.
```{r}
pheno2 <- pheno_1
pheno2$Y[pheno2$Y < 2] <- 0
pheno2
```
