---
title: "Katherine_VandePol_HW4"
author: "K. Vande Pol"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For homework 4, please pull the folder HW4 from the class_notesrepository on GitHub. It contains the pheno.txt* document to be used in this homework. From the pheno.txt file, obtain the following information:

Starting by reading the file into R:
```{r}
pheno <- read.table(file.choose("pheno.txt"), header = T, sep = "")
head(pheno)
```



1) How many unique observations are there in the “GENO” variable? (You haven’t used this function in class yet. Try to look for functions.)

```{r}
uniqueobs <- unique(pheno$GENO)
length(uniqueobs)
```
There are 485 unique observations of the "GENO" variable.



2) Save as a text file a subset of the pheno data set containing only range 1.
```{r}
phenorange1 <- pheno[pheno$range == 1,]
phenorange1
write.table(phenorange1, file = "phenorange1.txt")
```
phenorange1.txt now contains all of the data with the range variable = 1, without the data where range had other values.



3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.

First to find the range of the variable Y:
```{r}
range(pheno$Y, na.rm = T)
```
The range of Y is from minimum = 0.5250602 to maximum = 21.6188835.

To find the range of the variable M:
```{r}
range(pheno$M, na.rm = T)
```
The range of M is from minimum = 0.4502466 to maximum = 0.8779443.

To find the range of the variable h4:
```{r}
range(pheno$h4, na.rm = T)
```
The range of h4 is from minimum = 1.204 to maximum = 5.100.



4) What row contains the maximum value of “h4”?
```{r}
which.max(pheno$h4)
```
The maximum value is at 427th row of the variable h4.


5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).
```{r}
for (i in names(pheno[,7:12]))
  print(mean(pheno[,i], na.rm = T))
```
The mean of Y is 7.99446, of M is 0.7431033, for h1 is 0.2766773, for h2 is 1.248048, for h3 is 2.621128, and for h4 is 3.165503.


6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.
```{r}
pheno2 <- pheno
pheno2$Y[pheno2$Y < 2] <- 0
pheno2
```

The dataset pheno2 now contains a Y variable where all original values < 2 are set to 0.