---
title: "HW4_SarahWidener"
author: "Widener"
date: "9/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library("tidyverse")
callfile <- read.delim('~/HW4/HW4/pheno.txt')
#str(callfile)
```

1) How many unique observations are there in the “GENO” variable? (You haven’t used this function in class yet. Try to look for functions.)
```{r}
length(unique(callfile$GENO))
```

2) Save as a text file a subset of the pheno data set containing only range 1.
```{r}
subset(callfile, callfile$range == 1)
write.table(subset(callfile, callfile$range == 1), "Sarah")
```

3) The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables.
```{r}
(range_Y <- range(callfile$Y, na.rm = TRUE))
(range_M <- range(callfile$M, na.rm = TRUE))
(range_h4 <- range(callfile$h4, na.rm = TRUE))
```

4) What row contains the maximum value of “h4”?
```{r}
maxvalueh4 <-which(callfile$h4 == max(callfile$h4, na.rm = TRUE))
maxvalueh4
maxvaluerow <- which(callfile$h4 == maxvalueh4)
maxvaluerow
exactrow <- callfile[maxvaluerow,5]
exactrow
```
5) Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).
```{r}
abc <- callfile[,c("Y", "M", "h1", "h2","h3","h4")]
names<- variable.names(abc)
for (i in names) {
    x<- mean(abc[,i],na.rm = TRUE)
    print(x)
}
```
6) Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0.
```{r}
pheno2<-callfile
pheno2$Y[pheno2$Y<2]<-0
head(pheno2)
```



