---
title: "HW4"
author: "Winston Wong"
date: "9/19/2020"
output: html_document
---
pheno <- read.table("pheno.txt", header = TRUE)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 How many unique observations are there in the “GENO” variable?
```{r}
Unique_Genotypes <- length(unique(pheno$GENO))
Unique_Genotypes

```

#2 Save as a text file a subset of the pheno data set containing only range 1.
```{r}
Range1_Subset<- subset(pheno, range == 1)

write.table(Range1_Subset, file = "Range1_Subset.txt", sep = "\t",row.names = TRUE)

```

#3 The range (minimum and maximum value) of the “Y”, “M”, and “h4” variables
```{r}
Y_range <- range(pheno$Y, na.rm = T)
M_range <- range(pheno$M, na.rm = T)
h4_range <- range(pheno$h4, na.rm = T)

Y_range
M_range
h4_range
```

#4 What row contains the maximum value of “h4”?
```{r}
Max_h4 <- max(pheno$h4, na.rm = T)
Max_h4
is.numeric(Max_h4)

which(pheno$h4==Max_h4)

rowMax_h4 <- which(pheno$h4==Max_h4)
rowMax_h4

Variable_row_Max_h4 <- pheno[rowMax_h4, 5]
Variable_row_Max_h4
```

#5 Create a loop that prints the mean of the variables “Y”, “M”, “h1”, “h2”, “h3”, and “h4” (it must be done using a loop, e.g., for, while, or repeat).

```{r}

 means_of_columns <- vector("numeric", ncol(pheno[,7:12]))
    names(means_of_columns) <- names(pheno[,7:12])
    for (i in names(pheno[,7:12])) {
      means_of_columns[i] <- mean(pheno[[i]], na.rm = T)
    }
    means_of_columns

```

#6 Create a dataset called pheno2 and replace the values of “Y” that are smaller than 2 by 0
```{r}
pheno2 <- pheno
pheno2$Y[pheno2$Y<2] <- 0
pheno2

length(which(pheno2$Y == 0))
```

