---
title: "Homework 6"
author: "Beau Barber"
date: "10/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

___Using the data flights from the package nycflights13, do the following:___

```{r}
#Obtain data frame
install.packages("nycflights13")
flights <- nycflights13::flights
```


1- obtain the range, mean, and median of the variable “distance”.
```{r}
range_dist=range(flights$distance)
mean_dist=mean(flights$distance)
median_dist=median(flights$distance)
```

2- combine the airlines (also available from nycflights13) with the flights data in a single file.
```{r}
library(dplyr)
airlines <- nycflights13::airlines
#Use merge to combine the flights and airline data by the column they share, "carrier"
flights_airlines <- merge(airlines,flights,by="carrier")
head(flights_airlines)
```

3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}

aa_only <- flights_airlines %>%
  filter(carrier == "AA") %>%
  arrange(arr_delay)

save(aa_only,file = "aa_only.RData")

```


4- create boxplots of dep_delay by carrier (one boxplot for each carrier).

```{r}
library(ggplot2)
dep_delay_plot<-ggplot(flights_airlines, aes(x=carrier, y=(dep_delay)))+
  geom_boxplot(alpha = 0.5)+
  ggtitle("Departure Delays by Carrier")+
  labs(x="Carrier", y="Departure Delay (minutes)")
dep_delay_plot
```


5- find what is the month with the highest arr_delay. Show this in a graph.
```{r}
library(ggplot2)

avg_month_arr_delay <- group_by(flights, month)%>%
  summarise(avg_arr_delay=mean(arr_delay, na.rm=T))
head(avg_month_arr_delay)

#Max Average Arrival Delay row
max_month_arr_delay <- avg_month_arr_delay[which.max(avg_month_arr_delay$avg_arr_delay),]

p_plot<-ggplot(avg_month_arr_delay, aes(x=month, y=avg_arr_delay))+
  geom_col()+
  ggtitle("Average Arrival Delay each Month")+
  labs(x="Month", y="Average Arrival Delay (minutes)")#+
  scale_x_continuous(name="Months",limits=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
  #since I need month to be discrete and it is currently a continous variable



p_plot + geom_col(data = max_month_arr_delay, colour = 'red')

```
The month with the greatest average delay time is July.
