---
title: "Jadelyn_Stewart_HW6"
author: "Jadelyn Stewart"
date: "10/5/2020"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Using the data flights from the package nycflights13, do the following:

##1- obtain the range, mean, and median of the variable “distance”.

```{r}
library(tidyverse)
install.packages("nycflights13")
library(nycflights13)
flights <- flights

Q1 <- flights %>% summarise(range = range(distance, na.rm = TRUE),
                            mean = mean(distance, na.rm = TRUE),
                            median = median(distance, na.rm = TRUE))
Q1
#The range is 17 to 4983, mean is 1039.913, and median is 872
```

## 2- combine the airlines (also available from nycflights13) with the flights data in a single file.

```{r}
Q2 <- merge(flights, airlines, by = "carrier")
write.table(Q2, file = "Airlines and Flights Data.csv")
Q2
```

##3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}
Q3 <- Q2 %>% filter(carrier == "AA") %>% arrange(arr_delay)
write.table(Q3, file = "American Airlines by arr_delay.csv")
Q3
```
##4- create boxplots of dep_delay by carrier (one boxplot for each carrier).
```{r}
Q4 <- ggplot(flights, aes(carrier, dep_delay))+
  geom_boxplot(na.rm =TRUE)+
  ggtitle("Departure Delay by Carrier")+
  labs(x = "Carrier", y = "Departure Delay (mins)")
Q4
```

##5- find what is the month with the highest arr_delay. Show this in a graph.
```{r}
avg_arr <- group_by(flights, month) %>% summarize(avg_arr_delay = mean(arr_delay, na.rm=T))
avg_arr

highest_month <- avg_arr[which.max(avg_arr$avg_arr_delay),]
highest_month

#July has the highest arr_delay

Q5 = ggplot(avg_arr, aes(month, avg_arr_delay)) + geom_col() + 
    ggtitle("Average Arrival Delay by Month") + 
    labs(x = "Month", y = "Arrival Delays (mins)") + 
    scale_x_discrete(name="Month", limits = 
    c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))

Q5
```


