---
title: "Katherine_VandePol_HW6"
author: "K. Vande Pol"
date: "10/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the data flights from the package nycflights13, do the following:
1- obtain the range, mean, and median of the variable “distance”.

```{r}
library(tidyverse)
install.packages("nycflights13")
library(nycflights13)

head(flights)

dist <- flights %>% 
  summarise(
    range = range(distance, na.rm = TRUE),
    av = mean(distance, na.rm = TRUE),
    med = median(distance, na.rm = TRUE)
  )
dist

```
The range of the variable "distance" is 17 to 7983, the mean is 1039.913, and the median is 872.



2- combine the airlines (also available from nycflights13) with the flights data in a single file.

```{r}

combined <- merge(flights, airlines, by = "carrier")
write.table(combined, file = "Combined airlines and flights data.csv")

head(flights)
head(airlines)

head(combined)

```


3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.

```{r}

AA_only <- combined %>%
  filter(carrier == "AA") %>% 
  arrange(arr_delay)

write.table(AA_only, file = "AA_only.csv")

head(AA_only)

```


4- create boxplots of dep_delay by carrier (one boxplot for each carrier).

```{r}

dep_delay_box <- ggplot(flights, aes(x=carrier, y=dep_delay))+
  geom_boxplot(na.rm =TRUE)+
  ggtitle("Depart Delays by Carrier")+
  labs(x = "Carrier", y = "Depart Delay (min)")


dep_delay_box

```


5- find what is the month with the highest arr_delay. Show this in a graph.

Averaging each month's arrival delays:
```{r}

Av_arr_delay <- select(combined, arr_delay, month) %>%
  na.omit() %>%
  group_by(month) %>% 
  summarize(mean_arr_delay = mean(arr_delay))

Av_arr_delay

```


Plotting the average delays for each month as a column chart:
```{r}
p = ggplot(Av_arr_delay, aes(x = month, y = mean_arr_delay)) +
  geom_col(width = 0.8) +
  ggtitle('Arrival Delays by Month') + 
  labs(x = 'Month', y = 'Arrival Delays (min)') +
  scale_x_discrete(name="Month", limits = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))

p
```
The month with the greatest average delay time is July.

