---
title: "homework 6"
author: "Seda Ozer"
date: "10/4/2020"
output: html_document
---

```{r}
library(nycflights13)
library(dplyr)
library(tidyverse)
data(flights)
```
#1 obtain the range, mean, and median of the variable “distance”.
```{r}
range(flights$distance)
mean(flights$distance)
median(flights$distance)
```
#2- combine the airlines (also available from nycflights13) with the flights data in a single file.
```{r}
data("airlines")
combineddata<-merge(flights, airlines)
```
#3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}
Carrier_AA <- combineddata %>%
  filter(carrier == "AA") %>%
  arrange(arr_delay)

write.csv(Carrier_AA, file= "Carrier_AA.csv")
head(Carrier_AA)
```
#4- create boxplots of dep_delay by carrier (one boxplot for each carrier)
```{r}
library(ggplot2)
ggplot(combineddata, aes(x = factor(carrier), y = dep_delay)) +
  ggtitle("dep_delay by carrier")+
  geom_boxplot( 
    alpha=1,
    outlier.color="purple",
    outlier.size =1,
    na.rm=TRUE
)
```
#5-find what is the month with the highest arr_delay. Show this in a graph.
```{r}
ggplot(combineddata, aes(x = factor(month), y = arr_delay))+
         ggtitle("Arr_delays for each month")+
         geom_boxplot(
           alpha=1,
           outlier.color="red",
           outlier.size =1,
           na.rm=TRUE
)
```
```{r}
max(combineddata$arr_delay, na.rm = TRUE)
```
#The month with the highest arr_delay is January indicating the highest outlier.
I confirmed it via max function and found 1272 which is matching with boxplot.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
