Using the data flights from the package nycflights13, do the following:

```{r }
#install.packages("nycflights13")
library(nycflights13)
library(tidyverse)



data<-nycflights13::flights
```

1- obtain the range, mean, and median of the variable “distance”.

```{r }
range(data$distance, na.rm=T)

mean(data$distance, na.rm=T)

median(data$distance, na.rm=T)
```

2- combine the airlines (also available from nycflights13) with the flights data in a single file.

```{r }
data<-merge(data,(nycflights13::airlines))
```

if we need to save an actual file
write.csv(data,"HW6_Q2.csv")
3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.

```{r }
library(dplyr)

AAonly<-filter(data,data$carrier=="AA")
AAonly<-arrange(AAonly, desc(AAonly$arr_delay))


mean(AAonly$arr_delay,na.rm=T)
```

if we need to save an actual file
write.csv(AAonly,"HW6_Q3.csv")
4- create boxplots of dep_delay by carrier (one boxplot for each carrier).

```{r }
library(ggplot2)

ggplot(data,aes(x=carrier,y=dep_delay))+geom_boxplot()+(
  labs(title="Airline Departure Delay", x="Carrier",y="Departure Delay (mins)"))
```

5- find what is the month with the highest arr_delay. Show this in a graph.

```{r }
arrd<-group_by(data,month)

mpm<-arrd%>%summarize(mean=mean(arr_delay, na.rm=T))
```

keep the months in order on the x axis

```{r }
mpm$month<-factor(mpm$month,levels=c(1:12))

ggplot(mpm,aes(x=month,y=mean))+geom_bar(stat="identity")+
  labs(title="Arrival Delay By Month",x="Month",y="Arrival Delay (mins)")


library(knitr)
```

spin(,precious=T,doc="")
