---
title: "suzanna_storms_wk6"
author: "Suzanna Storms"
date: "10/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework should be submitted as an R Markdown document saved with your name (e.g., samuel_fernandes.Rmd) and pushed to the HW6 folder on theclass_notes repository in the IntroToR GitHub by next Monday. All code used to get the answers should be provided as R chunks.

Using the data flights from the package nycflights13, do the following:

1- obtain the range, mean, and median of the variable “distance”.
```{r}
library(nycflights13)
range_d <-range(flights$distance, ra.rm =TRUE)
median_d <-median(flights$distance, ra.rm=TRUE)
mean_d <- mean(flights$distance, ra.rm=TRUE)
range_d
median_d
mean_d
```
2- combine the airlines (also available from nycflights13) with the flights data in a single file.

```{r}
library(nycflights13)
library(dplyr)
name_flights<-full_join(flights,airlines)
```


3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}
AA_flights_arr <- arrange(name_flights,(arr_delay))

filter(AA_flights_arr, carrier == "AA")
```
4- create boxplots of dep_delay by carrier (one boxplot for each carrier).

```{r}
library(ggplot2)
dd_flights<- ggplot(data=name_flights, aes(x=carrier), y = dep_delay)
dd_flights+ geom_boxplot()+ ggtitle("Departure delays by carrier") + coord_flip()

```

5- find what is the month with the highest arr_delay. Show this in a graph.
```{r}
ad_plot <- ggplot(data= name_flights, aes(x= month), y=arr_delay)
ad_plot+ geom_histogram()+ ggtitle("Arrival delays by month")


```
