---
title: "WinstonWong_HW6"
author: "Winston Wong"
date: "10/3/2020"
output: html_document
---
```{r}
library(nycflights13)
data(flights)
as.data.frame(flights)
```

#1- obtain the range, mean, and median of the variable “distance”

```{r}
range_distance <- range(flights$distance)
range_distance

mean_distance <- mean(flights$distance)
mean_distance

median_distance <- median(flights$distance)
median_distance
```

#2- combine the airlines (also available from nycflights13) with the flights data in a single file.

```{r}
airlinesflights_merge <- merge(flights,airlines)
airlinesflights_merge
```

#3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}
#selecting variables "carrier" and "arr_delay"
amerAir <- select(airlinesflights_merge, carrier, arr_delay)

#selecting only "AA" in carrier variable
amerAir2 <- filter(amerAir, carrier == "AA")

#ordering by arr_delay
amerAir3 <- amerAir2[order(amerAir2$arr_delay),]

#saving the file
write.table(amerAir3, file = "amerAir3.txt", sep = "\t",row.names = TRUE)

#Alternatively we could use the pipe function
amerAir <- airlinesflights_merge %>% select(carrier, arr_delay) %>% filter(carrier == "AA") 
amerAir_ordered <- amerAir[order(amerAir$arr_delay),]

```

#4- create boxplots of dep_delay by carrier (one boxplot for each carrier).
```{r}

#Boxplot with R base
boxplot(dep_delay~carrier,data=airlinesflights_merge, main="dep_delay Data by Carrier",
   xlab="Carrier", ylab="dep_delay")
   
#Boxplot using ggplot2
as.factor(airlinesflights_merge$dep_delay)
ggplot(airlinesflights_merge, aes(x = carrier, y = dep_delay)) + 
  geom_boxplot()
```

#5 find what is the month with the highest arr_delay. Show this in a graph
```{r}
highest_arr_delay <- boxplot(arr_delay~month, data = airlinesflights_merge,
   main="arr_delay Data by Carrier",
   xlab="month", ylab="arr_delay")

#from the plot we can see that month 1 has the highest arr_delay value

plot(highest_arr_delay)

#checking work
max(airlinesflights_merge$arr_delay, na.rm = TRUE) #returns value of 1272 as max value in arr_delay
which(airlinesflights_merge$arr_delay == 1272,) #finding row number of value of 1272 which is 212861
airlinesflights_merge[212861,3] # returns month 1

sum(airlinesflights_merge$arr_delay == 1272, na.rm = TRUE) #checking number of instances of 1272, returns 1
```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
