---
title: "Ece Gulkirpik_HW6"
author: "Ece Gulkirpik"
date: "10/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#Using the data flights from the package nycflights13, do the following:
```{r}
#First of all, I installed the required packages for this assignment. Then,I wanted to display the summary statistics and structure of flights data

install.packages("nycflights13")
install.packages("dplyr")
install.packages("ggplot2")
library("nycflights13")
library("dplyr")
library("ggplot2")

summary(flights)
str(flights)
```

#1- obtain the range, mean, and median of the variable “distance”.
```{r}
var_dist <- flights %>%
  summarise(
    var_dist_range = range(distance, na.rm=TRUE),
    var_dist_mean = mean(distance, na.rm=TRUE),
    var_dist_median = median(distance, na.rm=TRUE)
  )
var_dist
```
#2- combine the airlines (also available from nycflights13) with the flights data in a single file
```{r}
combined_data <- merge(flights, airlines)
write.table(combined_data, file="flights_airlines_combined.txt")
```
#3- save a file including only American Airlines (Carrier = AA) ordered by arr_delay.
```{r}
airlines_american <- combined_data %>%
  filter(carrier == "AA") %>%
  arrange(arr_delay) %>%
  write.table("American_Airlines_Only.txt")
```
#4- create boxplots of dep_delay by carrier (one boxplot for each carrier).
```{r}
library(ggplot2)
plot_carrier_delay <- ggplot(combined_data, aes( x = carrier, y = dep_delay))+
  geom_boxplot(
    fill= "slateblue", 
    alpha=0.2,
    outlier.color="red",
    outlier.fill="red",
    outlier.size =3,
    na.rm=TRUE)+
  ggtitle('Departure Delay by Carrier')+
  labs( x = 'Carrier', y = 'Departure Delays (min)')
plot_carrier_delay
```

#5- find what is the month with the highest arr_delay. Show this in a graph.
```{r}
#First, I found which month had the highest arrival delay by using the following commands

library("nycflights13")
library("dplyr")
library("ggplot2")
arr_del_mo <- combined_data %>%
  select(arr_delay, month) %>%
  na.omit() %>%
  group_by(month) %>%
  summarise(avg_arr_delay = mean(arr_delay))
max_month <-which.max(arr_del_mo$avg_arr_delay)
max_month

#This showed that Month 7 - July has the highest average arrival delay. Next, I showed this in a bar graph.

plot_arrival_delay <- arr_del_mo %>%
  mutate(highlight_month = ifelse(month == '7', T,F))%>%
  ggplot(aes(x = month, y = avg_arr_delay))+
  geom_bar(aes(fill=highlight_month), stat = "identity", width=0.2, show.legend = FALSE)+
  scale_fill_manual(values = c('#595959', 'red'))+
  coord_flip()+
  labs(y = 'Average Arrival Delay(min)', x = 'Month')+
  scale_x_discrete(limits = c("Jan", "Feb", "Mar", "Apr", "May","Jun","Jul","Aug","Sept","Oct","Nov","Dec"))+
  ggtitle("Average Arrival Delay by Month")

plot_arrival_delay


```

