---
title: "suzanna_storms"
author: "Suzanna Storms"
date: "9/18/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(cache = TRUE)
```

## R Markdown

Question 1: uinque observations of GENO
```{r cars}

uGENO <- unique(pheno$GENO)
View(uGENO)
```
Question 2:
Save as a text file a subset of the pheno data set containing only range 1

```{r pressure, echo=FALSE}
library(tidyverse)
r1data <- pheno %>% filter(range < 2)
View(r1data)
```
Question 3:
the range of the Y, M and h4 variables
```{r}
rangeY <- range(pheno$Y, na.rm = TRUE)
rangeM <- range(pheno$M, na.rm = TRUE)
rangeh4 <- range(pheno$h4, na.rm = TRUE)
rangeY
rangeM
rangeh4
```
Question 4:
What row contains the maximum value of h4
```{r}
which.max(pheno$h4)
```

Question 5:
Create a loop that prints the mean of the variables Y, M, h1, h2, h4, and h4 (it must be done using a loop, eg. for, while, or repeat)


```{r}
i<-7
while(i<13){
  mc<-mean(pheno[,i], na.rm =TRUE)
  print(mc)
  i=i+1
}
#this is a function already in r that works as well
colMeans(pheno[,7:12], na.rm=TRUE)


```

Question 6: 
Create a dataset called pheno2 and replace the values of "Y" that are smaller than 2 by 0
# while
```{r}
pheno2 <- pheno
Y2<-replace(pheno$Y,pheno$Y<2,0)
pheno2[["Y"]] <- Y2
```


