---
title: "Homework7"
author: "Beau Barber"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Using the pheno.txt inside the HW4 folder, do the following:


Some libraries we might need
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(data.table)
library(pracma)
```

Get the data!
```{r}
pheno_dat<-read.table("https://raw.githubusercontent.com/IntroToR/class_notes/master/HW4/pheno.txt", head = T)
pheno_dat
```





## 1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)

___First we need to transform this data into long format___
```{r}
pheno_long <- pheno_dat %>% pivot_longer(names_to = "h_num", cols = c("h1","h2","h3","h4"), values_to="h_value")
head(pheno_long)
```

___Now to plot the histogram with h1, h2, h3, and h4___
```{r}
ggplot(data = pheno_long, aes(x=h_num,y=h_value, color=h_num))+
  geom_histogram(stat="identity")
```

___2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.___

Ope, looks like we have to use pivot_longer() again
```{r}
#Hello pivot_longer, my friend.
pheno_long2 <- pheno_dat %>% pivot_longer(names_to = "row_or_range", cols = c("row","range"))
#I need you to pivot again
#Because the data isn't in the right format
#and I can't plot the graphs right without that
head(pheno_long2)
```



```{r}
ggplot(data = pheno_long2, aes(y=Y,x=as.factor(value)))+
  geom_boxplot(fill="black",color="grey")+
  facet_wrap(~row_or_range,scales="free_x", ncol=2)+
  theme(axis.text.x = element_text(angle = 90, size=7))+
  xlab("Value")
#BUT HOW FOR THE LOVE STATS HOW DO YOU CHANGE THE X-AXIS INTERVAL FOR EACH
```


___3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).___

Assuming Y is the value we'll be taking the mean of
```{r}
pheno_dat %>% 
  group_by(TAXA) %>%
  summarise(taxa_mean=mean(Y)) %>%
  filter(taxa_mean>12) %>%
  ggplot(aes(x=as.factor(TAXA),y=taxa_mean))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=taxa_mean-1, ymax=taxa_mean+1), width=.5, position=position_dodge(.9))+ 
  theme(axis.text.x = element_text(angle = 90, size=9))+
  xlab("TAXA")

```


The folder HW7 contains one example graph for each question.