---
title: "Ece Gulkirpik_HMW 7"
author: "Ece Gulkirpik"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Using the pheno.txt inside the HW4 folder, do the following:

```{r}
#First, I started with getting the data and necessary packages for further questions:

library(ggplot2)
library(dplyr)
library(tidyverse)
pheno_data <- read.table("HW4/pheno.txt", header = TRUE)
pheno_data
```

#1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)

```{r}
#Transforming the "pheno_data" to a long format and creating a new data set called " pheno_data_long"

pheno_data_long <- pivot_longer(
  pheno_data,
  cols = c("h1","h2","h3","h4"),
  names_to = "h",
  values_to = "values"
)
pheno_data_long
```

```{r}
#Creating the histogram 
hist_var_h <- ggplot(pheno_data_long, aes(h, values)) + 
  geom_histogram(stat = "identity")+
  labs(x = "Measurement" , y = "Value")+
  ggtitle("Histogram of h")
hist_var_h
```

#2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.
```{r}
#First, I transformed the data again and created a new dataset called "pheno_data_rr"

pheno_data_rr <- pivot_longer(
  pheno_data, 
  cols = c("range", "row"), 
  names_to = "range_row", 
  values_to="value")
pheno_data_rr$value <- as.factor(pheno_data_rr$value)
pheno_data_rr

#Next, I made one boxplot containing one plot for row and another plot for range

boxp_var_y <- ggplot(pheno_data_rr,aes(value,Y))+
  geom_boxplot(alpha=0.2)+
  scale_x_discrete("value", seq(0,40,by=5))+
  facet_wrap(~range_row,scale="free_x", ncol=2)
boxp_var_y
```

#3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).

```{r}
#First, I grouped the data by TAXA variable, and calculated the mean of each TAXA. Finally, I filtered the TAXA with a mean greater than 12. I used pipe function to run all these function at once.

pheno_data_taxa <- pheno_data %>%
  group_by(TAXA) %>%
  summarise(mean=mean(Y, na.rm=T)) %>%
  filter(mean > 12)
pheno_data_taxa

#Next, I created the barplot with the error bars using geom_bar() and geom_errorbar() functions

taxa_bp <- ggplot(pheno_data_taxa, aes(TAXA, mean))+
  geom_bar(stat ="identity")+
  geom_errorbar(aes(ymin = mean-1, ymax = mean+1), width=0.5, colour="gray", size=2)+
  xlab("TAXA")+
  ylab("mean")+
  theme(axis.text.x=element_text(angle=90, size=9))
taxa_bp
  
```



