---
title: "Grace_Tan_HW7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
library(ggplot2)
library(dplyr)

pheno_data <-fread("https://raw.githubusercontent.com/IntroToR/class_notes/master/HW4/pheno.txt", header = T, sep = " ")
```

1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)

I decided to create two types of histograms with the data
```{r}
h_data <- pheno_data %>%
  select(h1, h2, h3, h4) %>%
  pivot_longer(cols = 1:4, names_to = "h", values_to = "values")

ggplot(h_data, aes(values, na.rm = TRUE)) +
  geom_histogram(bins = 50, aes(y = stat(count))) +
  ggtitle("Histogram of Values of h1, h2, h3, and h4") +
  labs(x = "Values", y = "Count")
```

The second kind is the same as the Example_1 image, and is the proper answer to this question.

```{r}
ggplot(h_data, aes(x=h, y=values))+
  geom_histogram(stat="identity")+
  ggtitle("Cumulative Values of h data")+
  labs(x = "h value", y="Cumulative Value")
```

2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.

```{r}
box_data <- pheno_data %>%
  select(range, row, Y) %>%
  pivot_longer(cols = 1:2, names_to = "row_range", values_to = "values")

ggplot(box_data, aes(x = values, y = Y)) +
  geom_boxplot(aes(group = cut_width(values, 1))) +
  facet_wrap(vars(row_range)) +
  labs(x="Values of Row and Range", y="Y Values")
```

3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).

```{r}
by_TAXA <- pheno_data %>% 
  group_by(TAXA)

TAXA_means <- by_TAXA %>%
  summarise(Y = mean(Y)) %>%
  filter(Y>12)

ggplot(TAXA_means, aes(x = TAXA, y = Y)) +
  geom_bar(stat = "identity") +
  ggtitle("Y Means By TAXA", "Means > 12") +
  theme(axis.text.x=element_text(angle = 40, hjust=1)) +
  labs(x = "Taxa", y = "Y Mean") +
  geom_errorbar(aes(ymin=Y-1, ymax=Y+1), width = 0.2)
  
```