---
title: "Jadelyn_Stewart_HW7"
author: "Jadelyn Stewart"
date: "10/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using the pheno.txt inside the HW4 folder, do the following:

```{r}
df1 <-read.table("https://raw.githubusercontent.com/IntroToR/class_notes/master/HW4/pheno.txt", head = T)
df1
```

##1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format) 

```{r}
library(tidyr)
df2 <- df1 %>% pivot_longer(names_to = "h_num", cols = c("h1","h2","h3","h4"), values_to="h_value")
df2

library(ggplot2)
ggplot(df2, aes(x = h_num, y = h_value)) +
  geom_histogram(stat = "identity") +
  labs(x = "Measurement", y = "Value")
```

##2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.
```{r}
df3 <- df1 %>% pivot_longer(names_to = "row_range_type", cols = c("row","range"))
df3

ggplot(data = df3, aes(x=as.factor(value), y=Y))+
  geom_boxplot(fill="black", color="grey")+
  facet_wrap(~row_range_type, scales="free", ncol=2)+
  xlab("Value")
```
##3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).
```{r}
library(dplyr)

mean <- df1 %>%
  group_by(TAXA) %>%
  summarise(mean = mean(Y, na.rm = T)) %>%
  filter(mean > 12)

ggplot(mean, aes(x = TAXA, y = mean)) + 
  geom_bar(position = position_dodge(), stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 1, ymax = mean + 1)) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Mean") + xlab("TAXA")
```

