---
title: "Katherine_VandePol_HW7"
author: "K. Vande Pol"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the pheno.txt inside the HW4 folder, do the following:

1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)

```{r}
HWdata <- read.table("https://raw.githubusercontent.com/IntroToR/class_notes/master/HW4/pheno.txt", header = T)
head(HWdata)

library(tidyr)
HWdata2 <- pivot_longer(HWdata, cols = c("h1", "h2", "h3", "h4"), names_to = "h")
head(HWdata2)

library(ggplot2)
ggplot(HWdata2, aes(x = h, y = value)) +
  geom_histogram(stat = "identity") +
  labs(x = "measurement", y = "value")

```



2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.

```{r}

HWdata3 <- pivot_longer(HWdata, row:range, names_to = "type", values_to = "values")
head(HWdata3)

ggplot(HWdata3, aes(x = as.factor(values), y = Y)) + 
  geom_boxplot(fill = "black", color = "gray") +
  ylab("Y") + xlab("values") +
  facet_wrap(~type, scales = "free")

```


3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).

```{r}
library(dplyr)

means <- HWdata %>%
  group_by(TAXA) %>%
  summarise(mean = mean(Y, na.rm = T)) %>%
  filter(mean > 12)

ggplot(means, aes(x = TAXA, y = mean)) + 
  geom_bar(position = position_dodge(), stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 1, ymax = mean + 1)) + 
  ylab("mean") + xlab("TAXA") +
  theme(axis.text.x = element_text(angle = 90))

```

