---
title: "HW 7"
author: "Stephen Schwartz"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pheno <- read.table("pheno.txt", header = T)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
```


#1. A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)
```{r}
Q1 <- pivot_longer(pheno, cols = c("h1", "h2", "h3", "h4"), names_to = "data", values_to = "values")

ggplot(Q1, aes(data, values)) +
  geom_histogram(stat = "identity")
```


#2. Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.
```{r}
Q2 <- pivot_longer(pheno, cols = c("range", "row"), names_to = "rangerow", values_to = "data")

ggplot(Q2, aes(as.factor(data), Y)) +
  geom_boxplot()+
  xlab("data")+
  facet_wrap(vars(rangerow), scales = "free")
```


#3. Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).
```{r}
Q3 <- pheno %>%
  group_by(TAXA) %>%
  summarise(averages = mean(Y, na.rm = T)) %>%
  filter(averages > 12)

ggplot(Q3, aes(TAXA, averages)) + 
  geom_col()+
  ylab("mean y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_errorbar(aes(ymin=averages-1,ymax=averages+1))

```


