---
title: "steven_krauklis_HW7"
author: "skrauklis"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#getting started w/ data and packages

```{r}
library(ggplot2)
library(tidyverse)
library(data.table)

pheno<-read.table("HW4/pheno.txt",sep="",header=T)
```

#1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)
```{r}
longp<-pheno%>%
  pivot_longer(names_to="htype",cols=c("h1","h2","h3","h4"),
               values_to="hval")

ggplot(longp,aes(x=htype, y=hval))+geom_histogram(stat="identity")
+labs(x="H type", Y="H Value")
```

#2) Using the variable "Y" as the response variable, create a boxplot '
#for each value of row and range. Make one plot for row
#and another plot for range using a facet option.
```{r}
lp2<-pivot_longer(pheno, cols=c("range", "row"),
                          names_to="range_or_row", values_to="rr_value")

ggplot(data = lp2, aes(x=as.factor(rr_value),y=Y))+
  geom_boxplot(fill="blue",color="black")+
  labs(x="Row/Range Value",y="Y")
```

#3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12
  #, and create a barplot with the error bar on top of it 
  #(use the value 1 as the standard error).
```{r}
Ymean<-pheno%>%
  group_by(TAXA)%>%
  summarize(mean=mean(Y,na.rm=T))%>%
  filter(mean>12)

#quick check
#filter(pheno,pheno$TAXA=="NSL102169")
#(7.809178+5.712128)/2

ggplot(Ymean,aes(x=TAXA,y=mean))+geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=mean-1,ymax=mean+1))+
  theme(axis.text.x=element_text(angle = 90, size=10))
```