---
title: "Sarah_Widener_W7.Rmd"
author: "Sarah_Widener_W7.Rmd"
date: "10/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#Using the pheno.txt inside the HW4 folder, do the following:
```{r}
library(dplyr)
library(tidyverse)
HW7data <- read.table("https://raw.githubusercontent.com/IntroToR/class_notes/master/HW4/pheno.txt", header = T)
head(HW7data)
library(tidyr)
HW7data2 <- pivot_longer(HW7data, cols = c("h1", "h2", "h3", "h4"), names_to = "h") 
head(HW7data2)
```

#1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)
```{r}
library(ggplot2)
ggplot(HW7data2, aes(x = h, y = value)) +
  geom_histogram(stat = "identity") +
  labs(x = "h_bin", y = "value")
```

#2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.
```{r}
HW7data3 <- pivot_longer(HW7data, row:range,names_to = "type", values_to = "values")
head(HW7data3)
ggplot(HW7data2) +
 aes(x = h, y = value) +
 geom_boxplot(fill = "#35b779") +
 theme_dark()
```

#3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).
```{r}
HW7data2%>%
  group_by(TAXA)%>%
  summarise_at(vars(Y), 
  list(name=mean)) 
mean2 <- HW7data2%>%
  group_by(TAXA)%>%
  summarise_at(vars(Y), 
  list(name=mean)) %>%
  filter(name>12)
n<-13
sd<- rep(1, each=n)
mean3<- mean2%>% add_column(sd)


finalplot <- ggplot(mean3)+
  geom_bar(aes(x = TAXA, y = name), stat="identity", fill="red", alpha=0.5)+
  geom_errorbar(aes(x=TAXA, ymin=name-sd, ymax=name+sd), colour="35b779",width=0.3, size=1)+
  xlab("TAXA")+
  ylab("Mean")+
 ggtitle("Average Y value by TAXA")
plot(finalplot)
```

