---
title: "WinstonWong_HW7.Rmd"
author: "Winston Wong"
date: "10/10/2020"
output: html_document
---
Downloading needed packages and data
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
pheno <- read.table("pheno.txt", header = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)
```{r}
#Converting pheno.txt into long format (2 ways)
#Way 1
pheno_long <- pivot_longer(pheno, cols=c("h1", "h2", "h3", "h4"), names_to="hnumber")

#Way 2
pheno_long <- gather(data = pheno, key = hnumber, value = measurement, h1:h4, factor_key = T)
refpheno_long <- filter(pheno_long[,9:10])

#Making a histogram using ggplot2
a <- ggplot(refpheno_long, aes(x= measurement)) +
  geom_histogram() +
  ylab("value") +
  ggtitle("Histogram of Measurements for h1, h1 ,h3, h4 ") +
  scale_fill_discrete(name = "Legend", labels = c("h1", "h2", "h3", "h4"))
a

```

2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.
```{r}
long_pheno <- pivot_longer(pheno, cols=c("range", "row"), names_to="Range_and_Row", values_to="Range_and_Row_Y")

ggplot(long_pheno, aes(x=as.factor(Range_and_Row_Y), y=Y))+
  geom_boxplot() +
  ggtitle("Measurements of Y by Range and Row Numbers")+
  labs(x="Range or Row", y="Y Measurement")+
  theme(axis.text.x=element_text(angle=90, hjust=1))+
  facet_wrap(~Range_and_Row, nrow=2, ncol=1, scale = "free_x")

```

3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error)
```{r}
#calculating mean of each Taxa and filtering with a mean
TAXA_mean <- pheno %>% group_by(TAXA) %>% summarize(mean = mean(Y, na.rm = T)) %>% filter(mean >12)

#Alternatively we could use:
TAXA_mean <- aggregate(x = pheno$Y, by = list(pheno$TAXA),FUN = mean) %>% filter(x > 12)
TAXA_mean

#Creating barplot with error bar
bp <- ggplot(TAXA_mean, aes(x = TAXA, y = mean)) + 
        geom_bar(stat = "identity") +
        theme(axis.text.x=element_text(angle=90, hjust=1)) +
        ggtitle("Average of Y for each TAXA") +
        geom_errorbar(aes(ymin=mean-1, ymax=mean+1), colour="black", width=.1)
```



