---
title: "Week 7 hw"
author: "Suzanna Storms"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the pheno.txt inside the HW4 folder, do the following:

1) A histogram with h1, h2, h3, and h4. (Tip: you need to transform this data in a long format)
```{r}
library(ggplot2)
library(tidyverse)
library(data.table)
pheno<-read.table("HW4/pheno.txt",sep="",header=T)

longpheno<-pheno%>%
  pivot_longer(names_to="htype",cols=c("h1","h2","h3","h4"),
               values_to="hval")
ggplot(longpheno,aes(x=htype, y=hval))+ geom_histogram(stat="identity")+labs(x="H type", y="H Value")

```


2) Using the variable "Y" as the response variable, create a boxplot for each value of row and range. Make one plot for row and another plot for range using a facet option.

```{r}
rrq2<-pivot_longer(pheno, cols=c("range", "row"),
                  names_to="range_row", values_to="data")
ggplot(data = rrq2, aes(x=as.factor(data),Y))+
  geom_boxplot()+
  labs(x="Row or Range",y="Y")
  facet_wrap(vars(range_row), scales = "free")

```



3) Calculate the mean of each TAXA, filter the TAXA with a mean greater than 12, and create a barplot with the error bar on top of it (use the value 1 as the standard error).

```{r}
mean_Y<-pheno%>%
  group_by(TAXA)%>%
  summarize(mean=mean(Y,na.rm=T))%>%
  filter(mean>12)
ggplot(mean_Y,aes(TAXA,mean))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=mean-1,ymax=mean+1))+
  theme(axis.text.x=element_text(angle = 90, size=9))


```

The folder HW7 contains one example graph for each question.
