---
title: "Ece_Gulkirpik_HW5"
author: "Ece Gulkirpik"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Using the R example data “mtcars”, do the following (when possible, try using the options available on the tidyverse package):

#1) Sort the dataset based on the variables “cyl” and “gear” (a single step, cyl is the first variable to be sorted and gear the second one).

install.packages("tidyverse")
install.packages("dplyr")

```{r}
data("mtcars")
mtcars
str(mtcars)
library(tidyverse)
mtcars_sorted <- arrange(mtcars,cyl,gear)
mtcars_sorted
```


#2) Create a variable called “model” with the names of the rows of this dataset.

```{r}
mtcars2 <- mutate(mtcars_sorted,model=row.names(mtcars_sorted))
mtcars2
```

#3) Replace the space of all the models by "_" (e.g., replace “Toyota Corona” by “Toyota_Corona”).

```{r}
gsub(" ", "_", mtcars2$model)
mtcars2
```
#4) Create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MGP greater than or equal 20)

```{r}
mpg_low <- filter(mtcars2, mpg < 20)
mpg_low
mpg_high <- filter(mtcars2, mpg >= 20)
mpg_high
```
#5) Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, " model“, and”gear“, create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.
```{r}
mtcars_new <- mtcars2 %>%
  arrange(hp) %>%
  select(cyl, hp, model, gear) %>%
  mutate(new_variable = cyl * gear)
mtcars_new
```

