---
title: "Grace_Tan_HW5"
date: "9/27/20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the R example data “mtcars”, do the following (when possible, try using the options available on the tidyverse package):

```{r}
library(tidyverse)
library(readr)
library(dplyr)
```

1) Sort the dataset based on the variables “cyl” and “gear” (a single step, cyl is the first variable to be sorted and gear the second one).

Using the arrange() function from the dplyr package within tidyverse!

```{r}
view(mtcars)

?arrange()
mtcars_ordered_cyl_gear <- arrange(mtcars, cyl, gear)
```


2) Create a variable called “model” with the names of the rows of this dataset.

```{r}
model <- rownames(mtcars)
view(model)
```


3) Replace the space of all the models by "_" (e.g., replace “Toyota Corona” by “Toyota_Corona”).

```{r}
model_subbed <- gsub(" ", "_", model)
view(model_subbed)
```

4) Create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MGP greater than or equal 20)

This is not a necessary step, but I arranged the mtcars data by increasing mpg, and then manually calculated which rows were less than 20 mpg, and which rows were greater than or equal to 20. 

```{r}
mtcars_ordered_mpg <- arrange(mtcars, mpg)
length(which(mtcars_ordered_mpg$mpg < 20))
length(which(mtcars_ordered_mpg$mpg >= 20))
```

Now I know that the first 18 rows should be grouped as mpg_low, and the last 14 rows should be grouped as mpg_high.  I will use the filter() function from the dplyr package within tidyverse to subset the data into two datasets.


```{r}
?filter()
mpg_low <- filter(mtcars_ordered_mpg, mpg < 20)
mpg_high <- filter(mtcars_ordered_mpg, mpg >= 20)

nrow(mpg_low) == 18
nrow(mpg_high) == 14
```


5) Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, "model“, and "gear", create a new variable named "new_variable" that is the multiplication of “cyl” x “gear”.

Model is not the name of a column under colnames(), so we must add it as the name of a column with the data created in question 3 using the add_column() function from dplyr.  I will also run all the functions individually to ensure they work before running them in a pipeline and viewing the final result, saved as mtcars_piped.

```{r}
?add_column

mtcars_model <- add_column(mtcars, model_subbed, .before = "mpg")
arrange(mtcars, hp)
select(mtcars, cyl, hp, gear)
mutate(mtcars, "new_variable" = cyl*gear)
```

Now that I have worked out which functions to use in the pipeline, and their proper arguments, I will construct my finals answer for question 5.

```{r}
mtcars_piped <- mtcars %>%
  arrange(hp) %>%
  add_column(model_subbed, .before = "mpg") %>%
  select(model_subbed, cyl, hp, gear) %>%
  mutate("new_variable" = cyl*gear)
  
view(mtcars_piped)
```




```
