---
title: "Hannah Demler HW5"
author: "Hannah Demler"
date: "9/22/2020"
output: html_document
---
**1**  
**Sort mtcars dataset based on the variables "cyl" and "gear"**  
```{r}
library(dplyr)
mtcars1<-mtcars[order(mtcars$cyl, mtcars$gear),]
mtcars1
```

**2**  
**Create variable called "model" with the names of the rows of this dataset**  
```{r}
model<-rownames(mtcars)
model
```

**3**  
**Replace the space of all the models by "_"**  
```{r}
row.names(mtcars)<-(gsub(" ","_",rownames(mtcars)))
```

**4**  
**Create two datasets, one with mpg_low and one with mpg_high**  
```{r}
mtcarslow<-mtcars[mtcars$mpg < "20",]
mtcarshigh<-mtcars[mtcars$mpg > "20",]
```

**5**  
**Using the pipe (%>%): sort by "hp", remove all columns different than "cyl", "hp", "model", and "gear", create a new variable named "new_variable" that is the multiplication of "cyl" x "gear"**  

```{r}
mtcars[order(mtcars$hp),] %>%
      select(-c(1,3,5,6,7,8,9,11)) %>%
      rownames_to_column('model')%>%
      mutate(new_variable= mtcars$cyl*mtcars$gear)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

