---
title: "Jadelyn_Stewart_HW5"
author: "Jadelyn Stewart"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Using the R example data “mtcars”, do the following (when possible, try using the options available on the tidyverse package):

##1) Sort the dataset based on the variables “cyl” and “gear” (a single step, cyl is the first variable to be sorted and gear the second one).



```{r}
install.packages('tidyverse','dplyr')
library(tidyverse)
library(dplyr)
data(mtcars)
df1 <- arrange(mtcars, mtcars$cyl, mtcars$gear)
df1
```

##2) Create a variable called “model” with the names of the rows of this dataset.

```{r}
model <- row.names(df1)
model
```

##3) Replace the space of all the models by "_" (e.g., replace “Toyota Corona” by “Toyota_Corona”).
```{r}
model <- (gsub(" ","_", model))
model
```
##4) Create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MGP greater than or equal 20)
```{r}
mpg_low <- filter(df1, df1$mpg < 20)
mpg_low

mpg_high <- filter(df1, df1$mpg >= 20)
mpg_high
```
##5) Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, " model“, and”gear“, create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.
```{r}
df1 = mutate(df1, model = model)
(q5 <- df1 %>% 
  arrange(df1, df1$hp) %>% 
  select(cyl, hp, model, gear) %>% 
  mutate(new_variable = cyl * gear))
```

