---
title: "Katherine_VandePol_HW5"
author: "K. Vande Pol"
date: "9/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the R example data “mtcars”, do the following (when possible, try using the options available on the tidyverse package):
1) Sort the dataset based on the variables “cyl” and “gear” (a single step, cyl is the first variable to be sorted and gear the second one).
```{r}
library(tidyverse)

head(mtcars)

HW5data <- arrange(mtcars, cyl, gear)

head(HW5data)

```



2) Create a variable called “model” with the names of the rows of this dataset.
```{r}

model <- row.names(HW5data)
model

```


3) Replace the space of all the models by "_" (e.g., replace “Toyota Corona” by “Toyota_Corona”).
```{r}

model <- (gsub(" ","_", model))
model

```


4) Create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MGP greater than or equal 20)
```{r}

mpg_low <- filter(HW5data, HW5data$mpg < 20)

mpg_low
  
mpg_high <- filter(HW5data, HW5data$mpg >= 20)
mpg_high

```


5) Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, " model“, and”gear“, create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.
```{r}

HW5data = mutate(HW5data, model = model)
(Problem5 <- HW5data %>% 
  arrange(hp) %>% 
  select(cyl, hp, model, gear) %>% 
  mutate(new_variable = cyl * gear))

```


