---
title: "suzanna_storms_wk5"
author: "Suzanna Storms"
date: "9/27/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(knitr)
library(tidyverse)
```



Use the "mtcars" package
question 1: sort hthe dtaset based on the variables "cyl" and "gear" (a single step, cyl is the first variable to be sorted and gear is the second one)

```{r cars}
cars<-mtcars
as_tibble(cars)
engine <- arrange(cars, cyl, gear)
```
Question 2: create a variable called "model" with the maes of the rows of the dataset.
```{r}
row.names.data.frame(cars)
model <-row.names.data.frame(cars)
```

Question 3: replace the space of all the models by "_".
```{r}
rownames_to_column(cars, var = "model")
cars2<- cars %>%
    mutate(model = str_replace(model, " ", "_"))
```
Question 4: create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MPG greater than or equal to 20)

```{r}
library(magrittr)
#%<>% #used to the PIPE to the original dataset
  
MPG_low<- cars2 %>%
  group_by(mpg) %>%
  filter(mpg < 20)
MPG_high<- cars2 %>%
  group_by(mpg) %>%
  filter(mpg >= 20)

```
Question 5:
Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, " model“, and”gear“, create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.

```{r}
Q5<- cars2 %>%
  select(cyl, model,hp, gear) %>%
  mutate(cyl *gear)
  
  

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
