---
title: "Winston_Wong_HW5.rmd"
author: "Winston Wong"
date: "9/28/2020"
output: html_document
---
```{r}
data(mtcars)
library(dplyr)
```

#1 Sort the dataset based on the variables “cyl” and “gear” (a single step, cyl is the first variable to be sorted and gear the second one).
```{r}
ordered <- arrange(mtcars, cyl, gear)
ordered
```

#2) Create a variable called “model” with the names of the rows of this dataset.
```{r}
mutant_mtcars <- mutate(mtcars, model = row.names(mtcars))
mutant_mtcars
```

#3 Replace the space of all the models by "_" (e.g., replace “Toyota Corona” by “Toyota_Corona”).
```{r}
mutant_mtcars$model <- gsub(" ", "_", mutant_mtcars$model)
```

#4 Create two datasets, one with mpg_low (by selecting cars with MPG less than 20) and mpg_high (by selecting cars with MGP greater than or equal 20)
```{r}
mpg_low <- subset(mutant_mtcars, mpg < 20,
select=everything(mutant_mtcars))
mpg_low

mpg_high <- subset(mutant_mtcars, mpg >= 20,
select=everything(mutant_mtcars))
mpg_high
```

#5 Using the pipe (%>%) command do all of these steps onto the mtcars dataset: sort by “hp”, remove all columns different than “cyl”, “hp”, " model“, and”gear“, create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.
```{r}
#sort by "hp"
sort_by_hp <- mtcars %>% arrange(hp)

# remove columns other than “cyl”, “hp”, " model“, and”gear“
remove_columns_different_than <- mutant_mtcars %>% within(rm(mpg, disp, drat, qsec, vs, am, carb))

# Create a new variable named”new_variable" that is the multiplication of “cyl” x “gear”.
mutant2_mtcars <- mutant_mtcars %>% mutate(new_variable = cyl*gear)
mutant2_mtcars
```

``